yield response function

### 2025-10-26 codes/R/visualized_results.qmd
modifications

models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF")

# 
plot_yield_response("NO_ANN", 1, global_models = models_to_compare)
plot_yield_response("NO_ANN", 5, global_models = models_to_compare)
plot_yield_response("NO_ANN", 10, global_models = models_to_compare)
plot_yield_response("DO_ANN", 1, global_models = models_to_compare)
plot_yield_response("DO_ANN", 5, global_models = models_to_compare)
plot_yield_response("DO_ANN", 10, global_models = models_to_compare)
plot_yield_response("SO_ANN", 1, global_models = models_to_compare)
plot_yield_response("SO_ANN", 5, global_models = models_to_compare)
plot_yield_response("SO_ANN", 10, global_models = models_to_compare)
plot_yield_response("NO_RF",  1, global_models = models_to_compare)
plot_yield_response("NO_RF",  5, global_models = models_to_compare)
plot_yield_response("NO_RF",  10, global_models = models_to_compare)
plot_yield_response("SO_RF",  1, global_models = models_to_compare)
plot_yield_response("SO_RF",  5, global_models = models_to_compare)
plot_yield_response("SO_RF",  10, global_models = models_to_compare)
```

@@

# ---- Packages ----
library(tidyverse)
library(fs)
library(here)
library(readr)
library(stringr)

@@

# --------
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

# number of fields
parse_n_fields <- function(folder) {
  f <- str_to_lower(folder)

  num <- str_match(f, "(_|\\b)(\\d+)_?fields?\\b")[,3]
  if (!is.na(num)) return(as.integer(num))

  words <- c("one"=1,"two"=2,"three"=3,"four"=4,"five"=5,
             "ten"=10,"twenty"=20,"thirty"=30,"forty"=40)
  word <- str_match(f, "(one|two|three|four|five|ten|twenty|thirty|forty)_?fields?")[,2]
  if (!is.na(word)) return(words[[word]])

@@

# --- model name mapping ---
parse_model <- function(folder) {
  f <- str_to_lower(folder)

  case_when(
    str_detect(f, "^cf_python")     ~ "CF_Python",
    str_detect(f, "^cf")            ~ "CF",
    str_detect(f, "^rann")          ~ "DO_ANN",
    str_detect(f, "^half_rann")     ~ "SO_ANN",
    str_detect(f, "^simple_ann")    ~ "No_ANN",
    str_detect(f, "^no_rf")         ~ "NO_RF",
    str_detect(f, "^halfrrf")       ~ "SO_RF",
    TRUE                            ~ NA_character_
  )
}

# read RDS, CSV
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF or CF_python csv
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# list files in dir 
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    keep(~ str_detect(path_file(.x),
                      regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                            ignore_case = TRUE))) %>%
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# extract sim id from filename
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

@@

# --- RMSE table ----
build_rmse_table <- function(base_dir = here("results"),
                             expected_rows = EXPECTED_ROWS) {

  dirs <- fs::dir_ls(base_dir, type = "directory", recurse = FALSE) %>%
    keep(~ str_detect(basename(.x),
                      regex("^(cf|cf_python|rann|half_rann|simple_ann|no_rf|halfrrf)",
                            ignore_case = TRUE)))

@@

@@

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(Model), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}

@@

# -- label mapping for models ---
model_names <- function(models) {
  map <- c(
    "Simple_ANN" = "No_ANN",
    "RANN"       = "DO_ANN",
    "HALFRANN"   = "SO_ANN",
    "HALFRRF"    = "SO_RF",
    "NO_RF"      = "NO_RF",
    "CF"         = "CF",
    "CF_Python"  = "CF_Python"
  )

>>>>

models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF", "CF")

# 
plot_yield_response("NO_ANN", 1, global_models = models_to_compare)
plot_yield_response("NO_ANN", 3, global_models = models_to_compare)
plot_yield_response("NO_ANN", 5, global_models = models_to_compare)
plot_yield_response("NO_ANN", 10, global_models = models_to_compare)
plot_yield_response("NO_ANN", 20, global_models = models_to_compare)
plot_yield_response("DO_ANN", 1, global_models = models_to_compare)
plot_yield_response("DO_ANN", 3, global_models = models_to_compare)
plot_yield_response("DO_ANN", 5, global_models = models_to_compare)
plot_yield_response("DO_ANN", 10, global_models = models_to_compare)
plot_yield_response("DO_ANN", 20, global_models = models_to_compare)
plot_yield_response("SO_ANN", 1, global_models = models_to_compare)
plot_yield_response("SO_ANN", 3, global_models = models_to_compare)
plot_yield_response("SO_ANN", 5, global_models = models_to_compare)
plot_yield_response("SO_ANN", 10, global_models = models_to_compare)
plot_yield_response("SO_ANN", 20, global_models = models_to_compare)
plot_yield_response("NO_RF",  1, global_models = models_to_compare)
plot_yield_response("NO_RF",  3, global_models = models_to_compare)
plot_yield_response("NO_RF",  5, global_models = models_to_compare)
plot_yield_response("NO_RF",  10, global_models = models_to_compare)
plot_yield_response("NO_RF",  20, global_models = models_to_compare)
plot_yield_response("SO_RF",  1, global_models = models_to_compare)
plot_yield_response("SO_RF",  3, global_models = models_to_compare)
plot_yield_response("SO_RF",  5, global_models = models_to_compare)
plot_yield_response("SO_RF",  10, global_models = models_to_compare)
plot_yield_response("SO_RF",  20, global_models = models_to_compare)
plot_yield_response("CF", 1, global_models = models_to_compare)
plot_yield_response("CF", 3, global_models = models_to_compare)
plot_yield_response("CF", 5, global_models = models_to_compare)
plot_yield_response("CF", 10, global_models = models_to_compare)
plot_yield_response("CF", 20, global_models = models_to_compare)

@@

library(tidyverse)
library(fs)
library(here)
library(readr)
library(stringr)

@@

# ---- RMSE function ----
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

# ---- number of fields parser ----
parse_n_fields <- function(folder) {
  f <- str_to_lower(folder)

  # numeric pattern
  num <- str_match(f, "(_|\\b)(\\d+)_?fields?\\b")[,3]
  if (!is.na(num)) return(as.integer(num))

  # text pattern
  words <- c("one"=1,"three"=3,"five"=5,"ten"=10,"twenty"=20)
  word <- str_match(f, "(one|three|five|ten|twenty)_?fields?")[,2]
  if (!is.na(word)) return(words[[word]])

@@

# ---- model name parser ----
parse_model <- function(folder) {
  f <- str_to_lower(folder)

  case_when(
    str_detect(f, "^cf")            ~ "CF",
    str_detect(f, "^rann")          ~ "DO_ANN",
    str_detect(f, "^half_rann")     ~ "SO_ANN",
    str_detect(f, "^simple_ann")    ~ "No_ANN",
    str_detect(f, "^no_rf")         ~ "NO_RF",
    str_detect(f, "^halfrrf")       ~ "SO_RF",
    TRUE                            ~ NA_character_
  )
}

# ---- read any file ----
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

# ---- read eonr file ----
read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# ---- list valid files ----
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    keep(~ str_detect(path_file(.x),
                      regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                            ignore_case = TRUE))) %>%
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# ---- extract sim id ----
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

# ---- build one row ----
row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

@@

# ---- build RMSE table ----
build_rmse_table <- function(base_dir = here("results"),
                             expected_rows = EXPECTED_ROWS) {

  dirs <- fs::dir_ls(base_dir, type = "directory", recurse = FALSE) %>%
    keep(~ str_detect(basename(.x),
                      regex("^(cf|cf_outcome|rann|half_rann|simple_ann|no_rf|halfrrf)",
                            ignore_case = TRUE)))

@@

    # only keep 1, 3, 5, 10, 20
    if (!n_f %in% c(1, 3, 5, 10, 20)) return(NULL)

@@

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(Model), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}
```

@@

rmse_tbl_raw <- readRDS(here("results", "rmse_tbl_raw.rds"))

# -- label mapping for models ---
model_names <- function(models) {
  map <- c(
    "Simple_ANN" = "No_ANN",
    "RANN"       = "DO_ANN",
    "HALFRANN"   = "SO_ANN",
    "HALFRRF"    = "SO_RF",
    "NO_RF"      = "NO_RF",
    "CF"         = "CF"
  )
